\vspace*{-5em}

My special thanks go to my supervisors Beno\^it Piranda, Julien Bourgeois and Seth Copen Goldstein. The supervision and support they have provided to me over the past years has truly helped the progression of my thesis and made my PhD journey a memorable, educational and enjoyable experience.

I express my sincere gratitude to my thesis committee members who have reviewed my dissertation. Their feedback and ideas have been absolutely invaluable.

Furthermore, I would also like to thank Abderahman Ait-Ali, my dear and faithful friend who is currently a PhD student at the KTH Royal Institute of Technology in Stockholm, for his availability and his constructive comments on my thesis work.

I would also like to thank my new office mate and recently PhD student, Pierre Thalamy, for having re-read the first three chapters of this manuscript and provided useful comments. I also thank him for his engineering work on VisibleSim as an intern in 2016. His code refactoring and test suite helped me to save time and focus on the design of my algorithms and simulation experiments. I wish him all the best for his PhD!

I would like to thank Catherine Maire for her proofreading that greatly improved the overall quality of this manuscript.

I am also thankful to the administrative staff of the University of Franche-Comté and to that of Carnegie Mellon University, in particular to Sylvie Klinkas, Dominique Ménétrier, Murielle Figuière, Marie-Thérèse Barthelet and Christina Contreras for their administrative support during my thesis.

Moreover, I would like to take this opportunity to thank the National Agency for Research (ANR) for funding my PhD research. This work is indeed part of the ANR - CO$^2$Dim project (contracts ANR-12-IS02-0004-01).

In addition, I would like to thank my enthusiastic office mates, Nicolas Boillot, Thadeu Knychala Tucci and, more recently, Pierre Thalamy, along with all the laboratory colleagues with whom I spent pleasant moments. Thadeu will definitely remain for ever in my mind as an incredible table soccer player (or not ;)! All joking aside, I wish him all the best for the end of his PhD!

%Last but foremost, my sincere thanks also go to all my family members and my beloved fiancée, Auréline Barrois, for their invaluable support and endless patience all along my studies. 
Mais avant tout, je tiens sincèrement à remercier ma famille et ma fiancée bien aimées pour leur soutien sans faille et leur patience sans limite durant mes années d'études. Je les remercie aussi pour leur relecture de mon manuscrit de thèse ainsi que leurs suggestions de corrections. Je remercie aussi tout particulièrement mes grands-parents d'être vaillamment venus assister à ma soutenance malgré leur grand âge et les nombreux kilomètres qui nous séparent. Je repense souvent aux cours de lecture de l'été 1996, chez mes grands-parents, pour combler mes lacunes à l'issue du cours préparatoire, aux nombreuses récitations de leçons le dimanche avec mes parents et aux coups de fil à ma tante pour discuter des cours d'histoire avant le brevet des collèges~! Que les années sont passées vite, merci à tous~!