{
	\vspace{-1.5cm}
	
	\renewcommand{\labelitemi}{}
	
	\newcommand{\pubEntry}[4]{
		#1. \textbf{{\color{ufcgreen}#2.}} \textit{#3,} #4. % title then journal/conf name and date.
	}
	%\addfontfeature{Color=blue}	
	\begin{itemize}
		\setlength{\itemsep}{0.925\baselineskip}			
		\item \pubEntry{André Naz, Benoît Piranda, Julien Bourgeois, and Seth Copen Goldstein}{A Time Synchronization Protocol for Large-Scale Distributed Embedded Systems with Low-Precision Clocks and Neighbor-to-Neighbor Communications}{In Journal of Network and Computer Applications (JNCA)}{accepted on December 2017. Elsevier}
		\item \pubEntry{André Naz, Benoît Piranda, Thadeu Tucci, Seth Copen Goldstein, and Julien Bourgeois}{Network Characterization of Lattice-based Modular Robots with Neighbor-to-Neighbor Communications}{In 2016 13th International Symposium on Distributed Autonomous Robotic Systems (DARS)}{pages 415 -- 429, London, UK, November 2016. Springer.}
		\item \pubEntry{André Naz, Benoît Piranda, Seth Copen Goldstein, and Julien Bourgeois}{A Time Synchronization Protocol for Modular Robots}{In PDP 2016, 24th Euromicro Int. Conf. on Parallel, Distributed, and Network-Based Processing}{pages 109 -- 118, Heraklion Crete, Greece, February 2016. IEEE. Core Rank: C. Acceptance Rate: 32\%}
		\item \pubEntry{André Naz, Benoît Piranda, Seth Copen Goldstein, and Julien Bourgeois}{A Distributed Self-Reconfiguration Algorithm for Cylindrical Lattice-based Modular Robots}{In 2016 IEEE 15th International Symposium on Network Computing and Applications (NCA)}{pages 254 -- 263, Cambridge, MA, USA, November 2016. IEEE. Core Rank: A. Acceptance Rate: 28\%}
		\item \pubEntry{André Naz, Benoît Piranda, Seth Copen Goldstein, and Julien Bourgeois}{Approximate-Centroid Election in Large-Scale Distributed Embedded Systems}{ In AINA 2016, 30th IEEE International Conference on Advanced Information Networking and Applications}{pages 548 -- 556, Crans-Montana, Switzerland, March 2016. IEEE. Core Rank: B. Acceptance Rate: 29\%}
		\item \pubEntry{Julien Bourgeois, Benoît Piranda, André Naz, Nicolas Boillot, Hakim Mabed, Dominique Dhoutaut, Thadeu Tucci, and Hicham Lakhlef}{Programmable Matter as a Cyber-Physical Conjugation}{In 2016 IEEE International Conference on Systems, Man, and Cybernetics (SMC 2016)}{pages 2942 -- 2947, Budapest, Hungary, October 2016. IEEE. Core Rank: B}
		\item \pubEntry{André Naz, Benoît Piranda, Seth Copen Goldstein, and Julien Bourgeois}{ABC-Center: Approximate-Center Election in Modular Robots}{In 2015 IEEE/RSJ International Conference on Intelligent Robots and Systems (IROS)}{pages 2951 -- 2957, Hamburg, Germany, September 2015. Core Rank: A. Acceptance Rate: 46\%}
	\end{itemize}
}